---
title: "Processing Nodes 4 VAIA"
author: "Francesco Pirotti"
date: "11/7/2020"
output:
  html_document: 
    toc: true 
    toc_float: true
---
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Some stuff on processing to share in the paper

## Aggregating data at nodes
Ok so we have nodes, and each 
A nice alternative for next time could be to have an hex grid, nice geometric advantages.   
Cannot convert our regular grid to hexagonal grid because nodes are in a regular grid, hexagonal requires offset at alternate rows.   

Grid is obviously not aligned to a regular latitude longitude grid as distance between nodes would decrease moving away from the equator.  

A custom projection Lambert Conformal Conical projection was designed (Marika Koukoula) with the following PROJ
  
```{r}
myproj<-"+proj=lcc +lat_1=45.827  +lat_2=45.827  +lat_0=45.827 +lon_0=11.625 +x_0=4000000 +y_0=2800000 +ellps=GRS80"
```

For each node a square with a small rotation was created.   The squares were uploaded to Google Earth Engine for part of the process.   



### Forest area and forest categories   
Italy is divided into regions and each region into provinces. We have a full cover of forest categories of two regions, Veneto and Friuli Venezia Giulia, and the province of Trento.   Categories slightly differ in classification, but discrepancies were merged into a single catogory.    


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
